\documentclass[11pt]{article}

\usepackage[cp1251]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{verbatim}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{enumerate}


\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\roots}{roots}
\DeclareMathOperator{\AG}{AG}
\DeclareMathOperator{\G}{G}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}


\begin{document}
\begin{center}
\textbf{The sizes of the gaps between roots after applying $(1-D)$ operator}\\
Stanislav Atanasov and Christopher Shriver
\end{center}

\section{Motivation}
Real-rooted polynomials (also referred to as \emph{hyperbolic polynomials}) are these polynomials who have only real roots. They have been object of many recent research~\cite{Borcea1, Borcea2, Spielman, Fisk} and their theory has had a significant progress in recent years~\cite{Borcea1, Borcea2}. A vast body of research deals with analyzing the linear transformation that map real-rooted polynomials. We say that a linear transformation $T:\mathbb{R}[x]\to\mathbb{R}[x]$ \emph{preserves real-rootedness}, or \emph{preserve hyperbolicity} if $T(f)$ is real-rooted, whenever $f$ is real-rooted. A seminal result by Borcea and Br$\ddot{a}$nden is the complete characterization of all linear transformations that preserve stability of multivariate polynomials, which in the univariate case amounts precisely to characterization of all linear real-rootedness preservers.

Despite its generality, the result by Borcea and Branden also has its limitations. It only asserts that a particular linear transformation $T:\mathbb{R}[x]\to\mathbb{R}[x]$ preserves real-rootedness but does not shed any light on where on the real line do the roots of $T(f)$ lie as some dependance on the roots of $f$. Instead of analyzing the position of the roots of $T(f)$, we investigate a related question. We observe the gaps between these roots after repeated application of a particular linear transformation which preserves real-rootedness.

In this paper we focus our attention to a particular linear operator that appears often in the recent papers by Marcus, Spielman and Srivastava~\cite{Spielman}. The linear operator $(1-D)$, where $D=\frac{d}{dx}$ is the differential operator, is known to preserve real-rootedness. Throughout the paper, we use $\lambda_{min}=\lambda_1\leq \cdots \leq \lambda_n = \lambda_{max}$ to denote the roots of the real rooted polynomial $P(x)$. Of particular interest for us are the \emph{average gap}, $\AG(P)$, of the polynomial $P(x)$, defined by $\AG(P)=\frac{\lambda_{max} - \lambda_{min}}{\deg P - 1}$, as well as the gaps between consecutive roots, which we will denote by $d_{i+1}(f) := \lambda_{i+1}(f) - \lambda_i(f)$. We establish the following two results

\begin{theorem}
please fill the exact formula
\end{theorem}

\begin{theorem}
Suppose there is a sequence $\{f_k\}_1^\infty$ of real-rooted polynomials defined by $f_k(x) = (1-D)^k f_0(x)$, where $f_0(x)$ is real-rooted of degree $m$. Then
\[\lim_{n\to\infty} \min_{i \in [m]} d_i (f_n) = \infty\]
holds true.
\end{theorem}
\section{Gaps}

\subsection{Average gap after many iterations}
 In this section, we show that the average gap approaches infinity when we apply $(1-D)$ operator many times. More precisely, for any sequence of real-rooted polynomials $\{f_k\}_1^{\infty}$, satisfying $f_{k+1}(x) = (1-D)^k f_{1}(x),$ it holds that $\lim_{n\to \infty} \AG(f_n) = \infty$. We prove this claim in two short steps.

 \begin{lemma}
 \label{degree two avg distance}
 Let $f(x)$ be a real-rooted quadratic polynomial with roots $\lambda_1 \leq \lambda_2$. Then $\AG((1-D)^n f) \to \infty$ as ${n\to \infty}$.
 \end{lemma}
\proof
Without loss of generality, we might assume that $f(x)$ is monic and write it as $x^2+ax+b$. The Vieta's formulas yield to $\AG(f) = \lambda_2 - \lambda_1 = \sqrt{a^2 - 4b}$. Introduce the sequence $\{f_k\}_{0}^\infty$ of polynomials given by $f_{k} (x)= (1-D)^k f_0(x)$ and $f_0(x) = f(x)$. Since $(1-D)$ preserves real-rootedness all $f_k(x)$ are real-rooted. In similar fashion, denote by $\lambda_1^{(k)} \leq \lambda_2^{(k)}$ the two real roots of $f_k(x)$. Using Vieta's formulas once again, we obtain
 \[\AG(f_k) = \sqrt{\AG(f_{k+1})^2 + 4}.~(*)\]

 Clearly, $\AG(f_k) > \AG(f_{k-1})$ for every $k\ge 1$ and so  $\lim \AG(f_k)$ exists (possibly, and as we shall see, infinite). Assume it is finite, say $l$. Taking the limit on both sides of $(*)$ leads to $l = \sqrt{l^2 + 4},$ which is evidently impossible. Hence, $\lim \AG(f_k) = \infty$. \qed

\begin{lemma}
\label{degree two - slowest movement}
Let $f(x)$ be a real-rooted polynomial with roots $\lambda_{min}=\lambda_1 < \lambda_2 \cdots \leq \lambda_{n-1} < \lambda_n = \lambda_{max}$. Then $AG((1-D)f) \geq AG[(1-D)f_{1,n}]$, where $f_{1,n} = (x - \lambda_{min})(x - \lambda_{max})$ is the quadratic polynomial with roots $\lambda_1$ and $\lambda_n$.
\end{lemma}
\proof Note that since $\lambda_{min}$ and $\lambda_{max}$ are simple roots, they are not roots of $(1-D)f$. Thus, the biggest and smallest root of $(1-D)f$ are the biggest and smallest root of $\frac{(1-D)f}{f}$. Let $x_0\in\mathbb{R}$ be biggest root of $(1-D)f(x)$. By the above remark, we have
\[\frac{(1-D)f}{f}(x_0) = 0 \Leftrightarrow \sum_{i=1}^n \frac{1}{x_0 - \lambda_i} = 1.\]

Note that $x_0 > \lambda_i,~\forall 1\leq i \leq n$ and so $\frac{1}{x_0 - \lambda_1} + \frac{1}{x_0 - \lambda_n} < \sum_{i=1}^n \frac{1}{x_0 - \lambda_i} = 1.$ In particular, since $\frac{(1-D)f_{1,n}}{f_{1,n}}(x) = \frac{1}{x - \lambda_1} + \frac{1}{x - \lambda_n}$ is monotonically decreasing function on $(\lambda_n, +\infty)$, the above inequality implies that the biggest root of $(1-D)f_{1,n}$ is smaller than $x_0$. Similarly, we prove that the smallest root of $(1-D)f_{1,n}$ is bigger than the smallest root of $(1-D)f$. Hence, the claim follows. \qed

Now, combining the two lemmata, we easily derive

\begin{proposition}For any sequence of real-rooted polynomials $\{f_k\}_0^{\infty}$ satisfying $f_{k+1}(x) = (1-D)^k f_{0}(x),$ we have $\lim_{n\to \infty} \AG(f_n) =  \infty.$
\end{proposition}
\label{avg gap goes to infty}
\proof Note that if $\lambda$ is a root of multiplicity $k$, then it is a root of multiplicity $k-1$ of $(1-D)f$. Therefore, after $deg f$ applications of the $(1-D)$ operator, we obtain a polynomial with simple roots. Therefore, we might, without loss of generality, assume that $f_0$ has simple roots. Otherwise, pick $f_{deg f_0}$ as a first element of the sequence. Suppose $\lambda_{min}=\lambda_1 < \lambda_2 < \cdots < \lambda_{n-1} < \lambda_n = \lambda_{max}$ are the roots of $f_0(x)$. Now repeated application of Lemma~\ref{degree two - slowest movement}, coupled with the observation that $\AG((1-D)[(x-a)(x-b)]) \geq \AG((1-D)[(x-c)(x-d)])$ whenever $c < a,~b<d$, leads to

\[\AG(f_k)= \AG((1-D)^kf_0)\geq\AG((1-D)^k [(x-\lambda_{min})(x-\lambda_{max})]),\]
for every $k\in\mathbb{N}$. Lemma~\ref{degree two avg distance} shows that, as $k$ grows, the right hand side goes to $\infty$, and thus so does the expression on the left. \qed

\subsection{The growth of $i$-th partial gap}
In the previous section, we concluded that the average gap increases without bound upon repeated applications of the $(1-D)$ operator. Proposition~\ref{avg gap goes to infty} asserts that the average gap goes to infinity. A natural question is to ask whether the same thing holds for each gap between two consecutive roots. As we shall see, this is indeed true.

\begin{lemma}
\label{minimal gap increases}
 Consider sequence of real-rooted polynomials $\{f_k\}_1^\infty$ given by $f_k(x) = (1-D)^k f_0(x)$. Then
 \[\min |\lambda^{(n+1)}_{i+1} - \lambda^{(n+1)}_i | \geq \min |\lambda^{(n)}_{i+1} - \lambda^{(n)}_i |\]
\end{lemma}
\proof Chris, please add it.
\qed

\begin{lemma}
Suppose $f_0(x)$ is a real-rooted polynomial with distinct roots $\lambda_1 < \lambda_2 < \cdots <\lambda_{n}$ such that $\lim_{m\to\infty} |\lambda_{i+1}[(1-D)^m f_0] - \lambda_{i}[(1-D)^m f_0]| =~\infty$ for every $ 1\leq i \leq n-1$. Suppose further $p_0(x)$ is a real polynomial having distinct roots $\mu_1 < \eta <\mu_2 < \cdots < \mu_{n-1},$ i.e., $p_0(x) = (x-\eta)f_0(x).$ Then we also have that
\[\lim_{n\to\infty} |\mu_{i+1}[(1-D)^m p_0] - \mu_{i}[(1-D)^m p_0]\big|=\infty,~\]
for every $1\leq i \leq n-1.$

\end{lemma}
\proof
We first show that 
\[|\mu_{2}[(1-D)^m p_0] - \mu_{1}[(1-D)^m p_0]| \geq |\lambda_{2}[(1-D)^m f_0] - \lambda_{1}[(1-D)^m f_0]|,~\forall m\in\mathbb{N}.\]


Let $\varepsilon > 0$. Now since $\lim_{m\to\infty} |\lambda_{2}[(1-D)^m f_0] - \lambda_{1}[(1-D)^m f_0]| = \infty$, there exist $M\in\mathbb{N}$ such that $|\mu_{2}[(1-D)^m p_0] - \mu_{1}[(1-D)^m p_0]| \geq \frac{1}{\varepsilon},~\forall m\geq M$. Pick $\delta > 0$ The roots of $(1-D)^{m+1} p_0$ satisfy
\[\sum_1^n \frac{1}{x-\mu^{(m)}_i} < \varepsilon \]


\qed

\begin{lemma}
\label{minimal gap shall approach}
 Consider a sequence $\{f_k\}_1^\infty$ of real-rooted polynomials defined by $f_k(x) = (1-D)^k f_0(x)$, where $f_0(x)$ is a real-rooted polynomial of degree~$m$. Suppose that for some $n\in\mathbb{N}$, there exist an index $1\leq i\leq m-1$ satisfying $d_{i-1}(f_n) \geq L$ and $d_{i+1}(f_n) \geq L$ for some real number $L$ bigger than $m$. Then we have that $\lambda^{(n+1)}_i$ is smaller than the smaller root of
 \[\frac{i-1}{L}+\frac{1}{x-\lambda^{(n)}_i} + \frac{1}{x-\lambda^{(n)}_{i+1}} = 1.\]
 Similarly, $\lambda^{(n+1)}_{i+1}$ is bounded from bellow by the bigger root of
 \[\frac{n-(i+1)}{L}+\frac{1}{x-\lambda^{(n)}_i} + \frac{1}{x-\lambda^{(n)}_{i+1}} = 1.\]
\end{lemma}
\proof
We only need to show the first bound since the second one can be proven analogously. We know that $\lambda^{(n+1)}_i\in(\lambda^{(n)}_i,\lambda^{(n)}_{i+1})$. Clearly, for every $x\in (\lambda^{(n)}_i,\lambda^{(n)}_{i+1})$ we have
\[\sum^{i-1}_{k=1} \frac{1}{x-\lambda^{(n)}_k} + \frac{1}{x-\lambda^{(n)}_i} + \frac{1}{x-\lambda^{(n)}_{i+1}} + \sum^n_{k = i+2} \frac{1}{x-\lambda^{(n)}_k} < \frac{i-1}{L}+\frac{1}{x-\lambda^{(n)}_i} + \frac{1}{x-\lambda^{(n)}_{i+1}}.\]
However, the inequality $L > m,$ implies $\frac{i-1}{L}<1$ and thus all roots of $\frac{i-1}{L}+\frac{1}{x-\lambda^{(n)}_i} + \frac{1}{x-\lambda^{(n)}_{i+1}} = 1$ are to the right of $\lambda^{(n)}_i$, or, more precisely, there is one root in each of the intervals $(\lambda^{(n)}_i,\lambda^{(n)}_{i+1})$ and $(\lambda^{(n)}_{i+1},+\infty)$. Now the claim follow since $\frac{1}{x-\lambda^{(n)}_i} + \frac{1}{x-\lambda^{(n)}_{i+1}}$ is monotonically decreasing in $(\lambda^{(n)}_i,\lambda^{(n)}_{i+1})$.
\qed

Now we have all the tools necessary for proving

\begin{proposition}
\label{minimal gap is unbounded}
Suppose there is a sequence $\{f_k\}_1^\infty$ of real-rooted polynomials defined by $f_k(x) = (1-D)^k f_0(x)$, where $f_0(x)$ is real-rooted of degree $m$. Then
\[\lim_{n\to\infty} \min_{i \in [m]} d_i (f_n) = \infty\]
holds true.
\end{proposition}
\proof Lemma~\ref{minimal gap increases} asserts the existence of such a limit. For the sake of contradiction, assume it is finite, say $l$.

We first analyze the behavior of the gaps between nonconsecutive roots of the $f_k(x)$ as $k$ increases. Applying Lemma~\ref{degree two avg distance} in combination with Lemma~\ref{degree two - slowest movement} easily show that all such gaps increase unboundedly and go to infinity. In particular, we have that
\[|\lambda^{(n)}_{i+2} - \lambda^{(n)}_i|\to \infty\text{~as~$n\to\infty$}\]
for every $1\leq i \leq n-2$.

Pick arbitrarily large number $L\in\mathbb{R}$ and some $\varepsilon > 0$. Then there exist a big enough $M$ so that for all $n\geq M$ we have that all double gaps satisfy $|\lambda^{(n)}_{i+2} - \lambda^{(n)}_i| > L + l$ and also $\min_{1\leq i \leq m} d_i (f_n)> l - \varepsilon$. Denote by $\{j_k\}_{n}^{\infty}$ the sequence of indexes such that $d_{j_n}:=\min_{1\leq i \leq m} d_i (f_n)$ for $n\geq M$. Since both distances $|\lambda^{(n)}_{j_n+1} - \lambda^{(n)}_{j_n-1}|$ and $|\lambda^{(n)}_{j_n+2} - \lambda^{(n)}_{j_n}|$ are bigger than $L + l$, we have that $d_{j_n-1}> L$ and $d_{j_n+1} > L$ for all $n\geq M$.

This means that we are in the conditions of Lemma~\ref{minimal gap shall approach}, which gives us a lower bound of the $d_{j_{n+1}}$ for every $n\geq M$. However, note that if $L$ is big enough, the quantities $\frac{i-1}{L}$ and $\frac{n-(i+1)}{L}$ from Lemma~\ref{minimal gap shall approach} become negligibly small. Therefore, by continuity, for every $\delta > 0$, there is $M$ such that the corresponding $L$ is big enough such that the roots $\lambda^{(n+1)}_{j_n}$ and $\lambda^{(n+1)}_{j_n}$ are both within $\frac{\delta}{2}$ of the roots $\mu^{(n)}_1,~\mu^{(n)}_2$ of $\frac{1}{x-\lambda^{(n)}_{j_n}} + \frac{1}{x-\lambda^{(n)}_{j_n+1}} = 1,~\forall n\geq M$. In particular, $|\lambda^{(n+1)}_{j_n+1} - \lambda^{(n+1)}_{j_n}|>|\mu^{(n)}_1 - \mu^{(n)}_2| - \delta,~\forall n\geq M$.

Recall that $|\mu^{(n)}_1 - \mu^{(n)}_2|= \sqrt{|\lambda^{(n)}_{j_n+1} - \lambda^{(n)}_{j_n}|^2 + 4}$. Now, since $\varepsilon > 0$ was picked arbitrarily, we might assume it is small enough to supply that $\sqrt{|\lambda^{(n)}_{j_n+1} - \lambda^{(n)}_{j_n}|^2 + 4} >l + \frac{\sqrt{l^2 -4} - l}{2}$ for all $n\ge M$ and so $|\mu^{(n)}_1 - \mu^{(n)}_2|\ge \frac{\sqrt{l^2 -4} - l}{2}$ for all $n\ge M$. Now, the choice of $\delta$ was independent of $\varepsilon$, so we pick $0< \delta <\frac{\sqrt{l^2 -4} - l}{2}$. This evidently leads to a contradiction with $l = \sup_{n\ge 0} (\min_{i \in [m]} d_i (f_n))$. Hence, the assumption that the limit $l$ is finite is wrong and thus $\lim_{n\to\infty} \min_{i \in [m]} d_i (f_n) = \infty$, as desired. \qed

\begin{thebibliography}{1}

  \bibitem{Borcea1} J. Borcea, and P. Branden {\em The Lee-Yang and Polya-Schur programs. I. Linear operators preserving stability.} Invent. Math. 177, pp. 541-569, 2009.

 \bibitem{Borcea2} J. Borcea, and P. Branden {\em The Lee-Yang and Polya-Schur programs. II. Theory of stable polynomials and applications} Comm. Pure Appl. Math. 62, pp. 1595-1631, 2009.

   \bibitem{Spielman} A. Marcus, D. Spielman, and N. Srivastava. {\em Ramanujan graphs and the solution of the Kadison-Singer problem.} Proceedings of ICM, Seoul, 2014.

   \bibitem{Fisk} S. Fisk. {\em Polynomials, roots, and interlacing.} available at \url{http://arxiv.org/abs/math/0612833}, 2008.
  \end{thebibliography}

\end{document}

